Tidy data structure to support exploration and modeling of temporal data

### 2018-11-20 main.Rmd
mostly done tweaking pipeline

There has been a long history of pipeline discussions and implementation centering around data in various aspects. The Extract, Transform, and Load (ETL) process is most commonly deployed pipeline in data science. However, building a sound data pipeline can be technically difficult, which involves many implementation decisions, including interface, input/output, functionality and so on, to be made.

@@

@viewing-pipeline proposed a viewing pipeline for data analysis in interactive statistical graphics and generalised a collection of elements required for the pipeline, which takes control of transformation from data to plot. @plumbing and @xie2014reactive implemented the data pipeline in interactive statistical software **plumbr** and **cranvas** respectively, using a reactive programming framework, in which user's interactions trigger a sequence of modules to update views accordingly.

The tidy data abstraction lays the foundation for data science modules including transformation, visualization and modelling. Each module frictionlessly communicates between each other. Because a desirable toolbox acquires tidy input, produces tidy output, and consequently chains a series of operations together to accomplish analytic tasks at hand.

What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the data analysis without concerning the details of computational implementation. A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.

>>>>

There has been a long history of pipeline discussions and implementation centering around data in various aspects. A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualising the process as it might be applied to a variety of problems. The Extract, Transform, and Load (ETL), from the data warehousing literature, outlines the workflow to prepare data for analysis, dates back to (XXX Caserta, Joe, 1965? REFS needed) can be considered a data pipeline. Building a data pipeline can be technically difficult, to make it sufficiently general for various data, with many implementation decisions on the interface, input and output objects and functionality. It is useful to articulate the data pipeline induced by new data tools.

@@

Buja et al. (1988) [-@viewing-pipeline] describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot. @xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **xgobi**, **ggobi**, **orca**, **plumbr** and **cranvas**, respectively. The pipeline is typically described with a one way flow, from data to plot. For interactive graphics, where all plots need to be updated when a user interacts with one plot, the events typically trigger the data pipeline to be run. @xie2014reactive uses a reactive programming framework, to implement the pipeline, in which user's interactions trigger a sequence of modules to update their views, that is, practically the same as running the data pipeline producing each plot.

The tidy data abstraction lays a pipeline infrastructure for data analysis modules, transformation, visualization and modelling. Each module communicates between each other, requiring tidy input, producing tidy output, and consequently chains a series of operations together to accomplish the analytic tasks at hand.

What is notable about an effective implementation of a data pipeline is that it coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies. A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.
